---
title: "Graduate Admission Analytics"
author: "Frimpong Atta Junior Osei & Ahmed Baruwa"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: united
    highlight: tango
    toc: true
    toc_float: 
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}

#knitr::opts_chunk$set(echo = TRUE)
#install.packages("r package", repos = "http://cran.us.r-project.org")
#library(tidyverse)
##install.packages("plotly")
#library(plotly)
#install.packages("corrplot")
#library(corrplot)
#install.packages("caTools")
#library(caTools)
#library(randomForest)
#library(modelr)
#library(MASS)
#library(dplyr)
#options(scipen = 100, max.print = 1e+06)
```

## Introduction

Graduate School application can be a very tedious process. Most candidates prepare their credentials with a very little knowledge about the details of the process. Even the most qualified and confident applicants worry about getting into graduate school. Unfortunately, graduate school admissions statistics tend to be more difficult to find than undergraduate acceptance rates. For this project, our main focus will be predicting the probability of a student getting admitted to graduate schools based on the following factors.

## Data

This dataset contains information or criteria for determining Postgraduate Admissions from an Indian perspective. This data was created by Mohan S. Acharya and can be found on Kaggle. The dataset is inspired by the UCLA Graduate Dataset with the aim of helping students get shortlisted in universities with based on their application materials. The predicted output gives them a fair idea about their chances for a particular university. The dataset contains several parameters which are considered important during the application for Masters Programs.

### Insights on Data
These are the variables that affect the data: 
-   GRE Scores ( out of 340 )
-   TOEFL Scores ( out of 120 )
-   University Rating ( out of 5 )
-   Statement of Purpose and Letter of Recommendation Strength ( out of 5 )
-   Undergraduate GPA ( out of 10 )
-   Research Experience ( either 0 or 1 )
-   Chance of Admit ( ranging from 0 to 1 )

```{r}
library(tidyverse)
adm_dataset <- read.csv("/Users/abaruwa/Desktop/PERSONAL/Spring\ 2022/Data\ Science/Project/newData/Admission.csv")
n=nrow(adm_dataset)
### 80/20 train/test split
library("MLmetrics") 
train_id= sample(n, n*0.8)
train_data = adm_dataset[train_id,]
test_data = adm_dataset[-train_id,]
attach(adm_dataset)
```

## Questions

Currently, it is totally understandable that the most critical factors to get into grad school, particularly for PhD programs depends on some or all the parameters mentioned earlier. Out of curiosity we plan to investigate the following questions:

-   Is the reputation of your previous institution a strong factor? 

-   Is research experience an important factor when thinking about getting into grad school? What would be the case for a fresh graduate with none? Can a catching Statement of Purpose (SOP) replace that?

-  Is there a correlation between GPA and Research ?

-   What is the pairwise correlation between all the variables ?

## Visualization

### Admission Chance Distribution

```{r}
ggplot(adm_dataset,aes(Chance_of_Admit)) + 
  geom_histogram(aes(fill=..count..),bins=40)
```

### Chance of admission by school rating

```{r}
ggplot(adm_dataset,aes(factor(University_Rating),
    Chance_of_Admit)) + geom_boxplot(aes(fill=University_Rating))
```
As we can see, students from highly ranked schools are more likely to be admitted if not all.

### GRE vs Research

```{r}
ggplot(adm_dataset, aes(GRE_Score, color=factor(Research)))+
  geom_density(alpha=0.5)+ggtitle("GRE vs Research Distribution")
```
Most students with more research experience are likely to get higher GRE Scores. 


### 3D Graph for chance of admission vs CGPA, TOEFL and GRE

```{r}
#install.packages("plotly")
library(plotly)
plot_ly(adm_dataset, x=~GRE_Score, y=~TOEFL_Score, z=~CGPA, color = ~Chance_of_Admit, 
        type="scatter3d", mode="markers") %>% 
  layout(scene = list(xaxis = list(title = "GRE Score"),
                      yaxis = list(title = "TOEFL Score"),
                      zaxis = list(title = "CGPA")))
```

From the 3D graph, we could see that chance of admission is higher when the CGPA, TOEFL and GRE score is higher.

### Correlation between variables

```{r}
library(corrplot)
corrplot(cor(adm_dataset), method = "number")
```

The more extreme the correlation coefficient (the closer to -1 or 1), the stronger the relationship. The positive correlation implies that the two variables under consideration vary in the same direction, i.e., if a variable increases the other one increases and if one decreases the other one decreases as well.

## Modeling

### Data Preparation - Splitting Data

```{r}
library("MLmetrics") 
train_id= sample(n, n*0.8)
train_data = adm_dataset[train_id,]
test_data = adm_dataset[-train_id,]
```

### Support Vector Machines

Let's try to do support vector machine modeling using Chance_of_Admit as the target variable.

```{r}
#install.packages("modelr")
library(e1071)
svm.fit <- svm(Chance_of_Admit~., data=train_data, type="eps", kernel="linear")
y_pred <- predict(svm.fit,test_data[,-c(8)])
R2_Score(y_pred, test_data$Chance_of_Admit)
```


### Random Forest Regression

Let's try to do random forest regression modeling using Chance_of_Admit as the target variable.
```{r}
library(randomForest)
rf.fit <- randomForest(Chance_of_Admit ~ ., data=train_data, ntree=100, keep.forest=TRUE, importance=TRUE)
y_pred <- predict(rf.fit, test_data[,-c(8)])
R2_Score(y_pred, test_data$Chance_of_Admit)
```

### XGBoost Regression

Let's try to do xgboost regression modeling using Chance_of_Admit as the target variable.

```{r}
library(xgboost)
xgb.fit <-xgboost(data= as.matrix(train_data[,-c(8)]), label=as.matrix(train_data$Chance_of_Admit), 
  nrounds = 1000, objective = "reg:squarederror", early_stopping_rounds = 3,
  max_depth = 6, eta = .25)
y_pred <- predict(xgb.fit, as.matrix(test_data[,-c(8)]))
R2_Score(y_pred, test_data$Chance_of_Admit)
```

## Analysis and Discussion
From the visualizations above the following three variables TOEFL_Score, GRE_Score, TOEFL_Score, have the highest correlation with the chances_of_admit of students. The p-value of SOP and GPA are more than 0.05, making them insignificant variables. This means they have no influence on the chance of getting admitted.

## Impact
If students can make use of the knowledge from these visualizations, students can tailor their applications around their strong credentials. The brightest students end up in the best of schools.  
